---
title: "An exciting interactive html file"
output: html_document
date: "2023-10-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Zooplankton data are variable
The structure of our zooplankton assemblages are variable over time.

We have lots of species from lots of monitoring stations.

It is often difficult to collapse this variability into a 2 dimensional ordination.

Maybe having an interactive 3d plot could offer some insight?

## Sites data only

```{r, echo=FALSE,message=FALSE,warning=FALSE}
### generate 3 dimensional NMDS plot for sites

#### load packages ####
ld_pkgs <- c("tidyverse","vegan","rgl")
vapply(ld_pkgs, library, logical(1L),
       character.only = TRUE, logical.return = TRUE);rm(ld_pkgs)
ord <- readRDS("data/out/ord3d.Rdata")
### load data ####
scores_site <- readRDS("data/out/scores_site3d.Rdata")
r3dDefaults$windowRect <- c(0,100, 1000, 1000) 
plot3d(x = scores_site$NMDS1,
       y = scores_site$NMDS2,
       z = scores_site$NMDS3,
       col=scores_site$RegionCol,
       type="s", size = 1,
       xlab = "NMDS1",ylab = "NMDS2",zlab = "NMDS3")

rglwidget(elementId = "nmds")
```

## Sites and species data

I'm trying to incorporate both sites and taxa into a single plot.
This version comes out a bit messy.  Can't see how to have semi-transparent text.
Potentially of some value?

```{r,echo=FALSE}
### generate 3 dimensional NMDS plot for species

### load data ####
scores_site <- readRDS("data/out/scores_site3d.Rdata")
# r3dDefaults$windowRect <- c(0,100, 1000, 1000) 
plot3d(x = scores_site$NMDS1,
       y = scores_site$NMDS2,
       z = scores_site$NMDS3,
       col=scores_site$RegionCol,
       type="s", size = 1,
       xlab = "NMDS1",ylab = "NMDS2",zlab = "NMDS3")
scores_species <- readRDS("data/out/scores_species3d.Rdata")
text3d(x = scores_species$NMDS1,
       y = scores_species$NMDS2,
       z = scores_species$NMDS3,
       scores_species$lb, size=1,
       cex = .575)

rglwidget(elementId = "nmds2",width=1000,height=1000)
```
Stress = `r round(ord$stress,3)`

# Ambitions/To do:
* Come up with approach to make text transparent
* Explore approaches to bring in ggplot style methods (non-trivial...)